Arbitrary Precision and Arbitrary $n$-Grams

### 2011-03-01 paper/paper.Rnw
test pp

	Each subsection contains two examples from the Shakespeare corpus and two
	from War and Peace.  These corpora were chosen because they were determined
	at rough glance to be the easiest to parse.

@@

		One of the two extensions we chose was to implement the probability and
		perplexity calculations using arbitrary precision arithmetic (we also
		implemented it using logarithmic arithmetic).  This can lead to slower
		processing, but	guarantees that there is no data loss until we move into
		processing floats during exponentiation.

@@

		Our hypothesis was that the arbitrary precision math would reveal a
		small amount of error in computing the perplexity logarithmically.  Our
		results seem to indicate that any such error is very small (differing
		only in the last digit in measuring the perplexity of whole corpora),
		but that arbitrary-precision mathematics is much slower than floating
		point mathematics with logs when finding the product of many numbers.
		See \ref{runtimes}:

		\begin{table}
		\caption{Running times with arbitrary and floating-point precision in
		seconds.  Average of 10 trials, trained on the Shakespeare training
		corpus, perplexity calculated for Shakespeare training corpus}
		\label{runtimes}
		\begin{tabular}{l | r r}
			Operation & Arbitrary Precision & Logarithmic \\
			\hline
			Trigram tallying & 136.936949968 & 137.023640156\\
			Probability computation & 63.5648961067 & 60.1038038731 \\
			Perplexity computation & \textbf{990.322752953} & 65.2168021202 \\
		\end{tabular}
		\end{table}

@@

>>>>

	%We trained our system on the \emph{Shakespeare} and \emph{War and Peace} corpora.

	Each subsection contains two examples from the Shakespeare corpus and two
	from War and Peace.  These corpora were chosen because they were determined
	at rough glance to be the easiest to parse. %and had the fewest problematic.

@@

		One of the two extensions we chose was to implement the probability and
		perplexity calculations using arbitrary precision arithmetic (we also
		implemented it using logarithmic arithmetic).  This can lead to slower
		processing, but in our profiling, most of the time was spent in
		dictionary manipulation rather than mathematics.  This approach
		guarantees that there is no data loss until we move into processing
		floats during exponentiation.

@@

		Our hypothesis was that the arbitrary precision math would reveal a
		small amount of error in computing the perplexity logarithmically.

@@

\begin{figure}
\begin{center}
<<echo=F,fig=T,height=9>>=
test.plot()
@
\end{center}
\caption{\label{fig:test}Perplexities of the test corpora by $n$-gram length and probability computation method for models trained on the respective training corpora.}
\end{figure}
